Whittaker Models for Real Groups

### 2024-04-15 whittaker.tex
Small edits

It is well known that a natural bijection between nilpotent $K$-orbits and nilpotent $G(\R)$-orbits can be defined in more elementary terms. This is the Kostant--Sekiguchi correspondence~\cite{}. Furthermore, it is a deep result of Schmid and Vilonen (the main result of~\cite{SV1}) that the wavefront set and the associated variety are exchanged by this correspondence. (As a historical aside, the Kostant--Sekiguchi correspondence was discovered on the basis that the correspondence $\AV(\pi) \leftrightarrow \WF(\pi)$ might be expressed in elementary terms.)

Therefore, a particular case of Schmid and Vilonen's theorem says: 
\begin{proposition} \label{Sekiguchi_result} The bijection (3)--(4) is induced by the Kostant--Sekiguchi correspondence.
\end{proposition} 
We shall do two things here. First, we will quote a result of \cite{AVAV} which makes this case of the Kostant--Sekiguchi correspondence completely explicit. Second, we shall give a new and elementary proof of Proposition~\ref{Sekiguchi_result}: this avoids the use of Schmid and Vilonen's difficult results, and instead uses the results above (in particular Proposition~\ref{JeffTasho_criterion}).

\subsubsection{The Sekiguchi correspondence for principal nilpotent orbits}

Let spell out the special case of the Sekiguchi correspondence which is of interest here. 
Suppose $E_\R\in \Op\cap i\g(\R)$. Then we may choose an $SL(2)$-triple $(H_\R,E_\R,F_\R)$ with $F_\R$  contained in $i\g(\R)$,
which implies $H_\R\in \g(\R)$. 
After conjugating by~$G(\R)$ we may further assume $\theta(E_\R)=-F_\R$.
The Sekiguchi correspondence takes the $G(\R)$-orbit of~$E_\R$ to the $K$-orbit of
$$
E_\theta=\frac12(-iE_\R-iF_\R+H_\R)\in \N\cap \s.
$$

@@

\subsubsection{A simple proof of Proposition~\ref{Sekiguchi_result}}

Consider the principal nilpotent element~$E_\pi$ defined in~\eqref{e:Epi}. We define another regular nilpotent element  $E_\R$ by using $E_\pi$ as $E_\theta$ in~\eqref{def_E_R}. We know that the associated variety of $\pi$ is the closure of $K \cdot E_\pi$ (Proposition~\ref{}), so the image of $\AV(\pi)$ by the Sekiguchi correspondence is the closure of the $G(\R)$-orbit of $E_\R$. We want to check that $\WF(\pi)$

@@

Now, we know that $\Kostant{E_\R}=E_\R+\mathrm{Cent}_\g(F_\R)$ contains $E_\R-F_\R = H_\theta$. In general the orbits $\mathcal{O}_\pi$ and $G(\R) \cdot H_\theta$ are disjoint. Nevertheless $H_{\lambda}$ and $H_\theta=[E_\theta, F_\theta]$ are in the same Weyl chamber. Therefore the result follows from:

\begin{proposition}\label{prop:AC_chamber} Let $\mathcal{C} \subset \t$ be a large open Weyl chamber.
If $H$ and $H'$ are two points of~$\mathcal{C}$ and if $\mathcal{O}, \mathcal{O}'$ are the   $G(\R)$-orbits of $H$ and $H'$, then $\AC(\mathcal{O})=\AC(\mathcal{O}')$.
\end{proposition}

@@

\begin{lemma}\label{lem:Kost_and_AC} Let $\mathcal{O}$ be a regular semisimple orbit and  let $X$ be a regular nilpotent element. Then $\Kostant{X}$ meets~$\mathcal{O}$ if and only if $X$ is in the asymptotic cone~$\AC(\mathcal{O})$. 
\end{lemma}
\begin{proof} \textcolor{red}{See if we spell it all out (keeping in mind that half of it is in the paper with Tasho, but should probably be included for convenience since part of the arguments are similar). Or maybe it's actually possible to give just a reference to Fresse--Mehdi.} \end{proof}

@@

Let us prove Proposition~\ref{prop:AC_chamber}.

Given an element $H$ of $\mathcal{C}$ we shall write $\mathcal{O}_H$ for the $G(\R)$-orbit of~$H$. For any regular nilpotent orbit~$\Omega$ let $E^{\mathcal{C}}_\Omega$  denote the set of $H \in \mathcal{C}$ such that $\AC(\mathcal{O}_H)$ contains $\Omega$; byLemma~\ref{lem:AC_containment}  we know that   $E^{\mathcal{C}}_\Omega$ is an open subset of~$\mathcal{C}$.

@@

\subsection{Where to put this, if we keep it?}

There are several ways to describe this bijection.\footnote{AA 23-II-2024: I think we should mention the version with the asymptotic cone of the orbit of~$H_\pi$, which is extremely clean and concrete. } We choose to compose the maps (1)$\leftrightarrow$(3)$\leftrightarrow$(4).
Here is the conclusion.

>>>>

It is well known that a natural bijection between nilpotent $K$-orbits and nilpotent $G(\R)$-orbits can be defined in more elementary terms. This is the Kostant--Sekiguchi correspondence~\cite{sekiguchi}. Furthermore, it is a deep theorem of Schmid and Vilonen (the main result of~\cite{SV1}) that the wavefront set and the associated variety are exchanged by this correspondence. (As a historical aside, the Kostant--Sekiguchi correspondence was discovered on the basis that the correspondence $\AV(\pi) \leftrightarrow \WF(\pi)$ might be expressed in elementary terms.)

Therefore, a particular case of Schmid and Vilonen's theorem says: 
\begin{proposition} \label{Sekiguchi_result} The bijection (3)--(4) is induced by the Kostant--Sekiguchi correspondence.
\end{proposition} 
We shall do two things here. First, we will quote a result of \cite{AVAV} which makes this case of the Kostant--Sekiguchi correspondence completely explicit. Second, we shall give a new and elementary proof of Proposition~\ref{Sekiguchi_result}: this avoids the use of Schmid and Vilonen's difficult results, and instead uses the results above (in particular Proposition~\ref{JeffTasho_criterion}) as the main ingredients.

\subsubsection{The Sekiguchi correspondence for principal nilpotent orbits}\label{sec:concrete_sek}
\footnote{AA 2024-IV-15: should correct the formulas here.}
Let us spell out the special case of the Sekiguchi correspondence which is of interest here. 
Suppose $E_\R\in \Op\cap i\g(\R)$. Then we may choose an $SL(2)$-triple $(H_\R,E_\R,F_\R)$ with $F_\R$  contained in $i\g(\R)$,
which implies $H_\R\in \g(\R)$. 
After conjugating by~$G(\R)$ we may further assume $\theta(E_\R)=-F_\R$.
The Sekiguchi correspondence takes the $G(\R)$-orbit of~$E_\R$ to the $K$-orbit of
$$
E_\theta=\frac12(-iE_\R-iF_\R+H_\R)\in \N\cap \s.
$$

@@

\subsubsection{A simple proof of Proposition~\ref{Sekiguchi_result}}\label{sec:sekiguchi_proof}

Consider the principal nilpotent element~$E_\pi$ defined in~\eqref{e:Epi}. We define another regular nilpotent element  $E_\R$ by using $E_\pi$ as $E_\theta$ in~\eqref{def_E_R}. We know that the associated variety of $\pi$ is the closure of $K \cdot E_\pi$ (Lemma~\ref{l:pi_to_av}), so the image of $\AV(\pi)$ by the Sekiguchi correspondence is the closure of the $G(\R)$-orbit of $E_\R$. We want to check that $\WF(\pi)$

@@

Now, we know that $\Kostant{E_\R}=E_\R+\mathrm{Cent}_\g(F_\R)$ contains $E_\R-F_\R = H_\theta$\footnote{AA 2024-IV-15: insert the correct version once we've checked the Sekiguchi formulas in the previous paragraph. }
. In general the orbits $\mathcal{O}_\pi$ and $G(\R) \cdot H_\theta$ are disjoint. Nevertheless $H_{\lambda}$ and $H_\theta=[E_\theta, F_\theta]$ are in the same Weyl chamber. Therefore the result follows from:

\begin{proposition}\label{prop:AC_chamber} Let $\mathcal{C} \subset \t$ be a large open Weyl chamber.
If $H$ and $H'$ are  points of~$\mathcal{C}$ and $\mathcal{O}, \mathcal{O}'$ are the   $G(\R)$-orbits of $H$ and $H'$, then $\AC(\mathcal{O})=\AC(\mathcal{O}')$.
\end{proposition}

@@

\begin{lemma}\label{lem:Kost_and_AC} Let $\mathcal{O}$ be a regular semisimple orbit and  let $X$ be a regular nilpotent element. Then $\Kostant{X}$ meets~$\mathcal{O}$ if and only if $X$ is in the asymptotic cone~$\AC(\mathcal{O})$. 
\end{lemma}
\begin{proof} See if we spell it  out (one direction is in Adams--Kaletha, but parts of the arguments are similar so it might be more convenient include both directions...). Maybe it's actually possible just refer to Fresse--Mehdi with pointers. \end{proof}

@@

Let us prove Proposition~\ref{prop:AC_chamber}. Given an element $H$ of $\mathcal{C}$ we shall write $\mathcal{O}_H$ for the $G(\R)$-orbit of~$H$. For any regular nilpotent orbit~$\Omega$ let $E^{\mathcal{C}}_\Omega$  denote the set of $H \in \mathcal{C}$ such that $\AC(\mathcal{O}_H)$ contains $\Omega$; by Lemma~\ref{lem:AC_containment}  we know that   $E^{\mathcal{C}}_\Omega$ is an open subset of~$\mathcal{C}$.

@@

\subsubsection{Where to put this, if we keep it?}

As we saw the correspondence $\pi \leftrightarrow \WF(\pi)$ can be described using the asymptotic cone of the semisimple orbit of the Harish-Chandra parameter. We used this in the proof of Proposition~\ref{Sekiguchi_result}. Now, the results of Sections~\ref{sec:concrete_sek} and~\ref{sec:sekiguchi_proof} provide another description of $\WF(\pi)$ from the Harish-Chandra parameter: we can compose the maps (1)$\leftrightarrow$(3)$\leftrightarrow$(4). This gives the following rather concrete conclusion.
